---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# countryscales

<!-- badges: start -->
<!-- badges: end -->

`countryscales` extends [`scales`](https://scales.r-lib.org) and 
[`ggplot2`](https://ggplot2.tidyverse.org) by providing functions to make 
it easy to display numbers or label axis text on positional scale in decimal 
format, as percentages or currencies using country- or locale-specific style 
conventions.

## Installation

You can install the development version of `countryscales` from GitHub using:

``` r
remotes::install_github("trekonom/countryscales")
```


## Note on supported locales

`countryscales` uses data on locale-specific numbering formats from the Common Locale 
Data Repository ([CLDA](https://en.wikipedia.org/wiki/Common_Locale_Data_Repository)) 
provided for easy use in R by the [`i18n`](https://rich-iannone.github.io/i18n/) package. 
Currently `countryscales` supports 552 of the 574 locales listed in `i18n`.  Not 
supported are locales which deviate from the international norm to group digits 
by threes, i.e. locales using the 
[Indian numbering system](https://en.wikipedia.org/wiki/Indian_numbering_system) 
which

> groups the rightmost three digits together (until the hundreds place), and thereafter groups by sets of two digits.

## Usage

The most common use case for countryscales is to customise the appearance of 
axis and legend labels or format numbers added as labels to a plot using 
country-specific style conventions.

Here is first example showing how 1 million USD are formatted according to style 
conventions in the G7 countries:

```{r locale-g7}
library(countryscales)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)

g7 <- data.frame(
  country = c("Canada", "France", "Germany", "Italy", "Japan", "the United Kingdom", "the United States"),
  locale = c("en-CA", "fr-FR", "de-DE", "it-IT", "ja-JP", "en-GB", "en-US")
) |>
  mutate(
    value = purrr::map_chr(locale, ~ label_currency_locale(locale = .x, currency = "USD")(1e6))
  )

ggplot(g7, aes(x = factor(1), y = country)) +
  geom_label(
    aes(label = paste(value, "in", country)),
    fontface = "bold"
  ) +
  labs(title = "In the G7 countries, 1 million USD is written as")
```

```{r base}
base <- gapminder15 |>
  count(region, wt = pop) |>
  ggplot(
    aes(n, reorder(region, n),
      fill = region
    )
  ) +
  scale_fill_brewer(palette = "Dark2") +
  geom_col(width = .6) +
  theme_minimal() +
  labs(
    x = NULL, y = NULL,
    title = "Population by World Regions in 2015",
    subtitle = "Default ggplot2"
  ) +
  guides(fill = "none")
```

As a second example consider formatting a chart according to German style conventions 
where a dot (`.`) is used as the big mark.

Using `countryscales` this could be achieved using the `scale_x/y_xx_locale` 
and  `label_xxx_locale` family of functions like so:

```{r locale-de}
base +
  geom_label(
    aes(
      label = label_number_locale(
        locale = "de-DE",
        accuracy = 1000
      )(n)
    ),
    hjust = 1,
    fill = NA,
    label.size = NA,
    color = "white"
  ) +
  scale_x_number_locale(
    locale = "de-DE",
    expand = expansion(mult = c(0, .05))
  ) +
  labs(subtitle = "... using German style conventions.")
```

Besides the `_locale` family of functions countryscales provides 
convenience functions for some common locales, e.g. to format the plot using 
Swiss style conventions where a right single quote (`r knitr::asis_output("\u2019")`) = 
\\u2019) is used as the big mark, you could achieve the same result using `label_number_ch` 
and `scale_x_number_ch` like so:

```{r locale-ch}
base +
  geom_label(
    aes(
      label = label_number_ch(accuracy = 1000)(n)
    ),
    hjust = 1,
    fill = NA,
    label.size = NA,
    color = "white"
  ) +
  scale_x_number_ch(
    expand = expansion(mult = c(0, .05))
  ) +
  labs(subtitle = "... using Swiss style conventions.")
```
