---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# countryscales

<!-- badges: start -->
<!-- badges: end -->

`countryscales` extends the [`scales`](https://scales.r-lib.org) package to make 
it easy to display numbers according to country-specific style conventions. 
Additionally `countryscales` extends [`ggplot2`](https://ggplot2.tidyverse.org) 
by providing a set of positional scale functions to make it easy to style 
axes displaying numbers, percentages or currencies.

## Installation

You can install the development version of `countryscales` from GitHub using:

``` r
remotes::install_github("trekonom/countryscales")
```

## Usage

The most common use case for countryscales is to customise the appearance of 
axis and legend labels or format numbers added as labels to a plot using 
country-specific style conventions.

As a first example consider formatting a chart according to style conventions 
used in Germany and several other European countries where a dot (`.`) is used 
as the big mark or grouping mark or thousands seperator and a comma (`,`) as the 
decimal mark.

Using `countryscales` this could be achieved using the `scale_x/y_xxx_locale` 
and  `label_xxx_locale` family of functions like so:

```{r}
library(countryscales)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)

base <- gapminder15 |>
  count(region, wt = pop) |>
  ggplot(aes(n, reorder(region, n), fill = region)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_col(width = .6) +
  theme_minimal() +
  labs(x = NULL, y = NULL, title = "Population by World Regions in 2015") +
  guides(fill = "none")

base +
  geom_label(aes(label = label_number_locale(locale = "de-DE", accuracy = 1000)(n)),
    hjust = 1, fill = NA, label.size = NA, color = "white"
  ) +
  scale_x_number_locale(locale = "de-DE", expand = expansion(mult = c(0, .05))) +
  labs(subtitle = "... using German style conventions.")
```

Besides the `_locale` family of functions countryscales provides 
convenience functions for most common locales, e.g. to format the plot using 
Swiss style conventions you could achieve the same result using `label_number_ch` 
and `scale_x_number_ch` like so:

```{r}
base +
  geom_label(aes(label = label_number_ch(accuracy = 1000)(n)),
    hjust = 1, fill = NA, label.size = NA, color = "white"
  ) +
  scale_x_number_ch(expand = expansion(mult = c(0, .05))) +
  labs(subtitle = "... using Swiss style conventions.")
```
