---
title: "Background"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Background}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `scales` package provides a family of easy to use functions 
like `scales::label_number` or `scales::label_percent` to format numbers in 
decimal format, as percentages or currencies, which could also be 
passed to the `labels` argument of `ggplot2`s family of scale functions to 
format axes or legends.

However, let's say you have to prepare a report or charts where you have to use 
country specific style conventions to display numbers which differ from the 
default anglo-american style.
  
While that could be achieved with the `scales` package as well, in practice 
formatting numbers and styling axes quickly becomes 
cumbersome and annoying as you always have to

* set the decimal and grouping marks (if deviating from the defaults) so you end up with
  ```
  scales::label_number(big.mark = ".", decimal.mark = ",")(x)
  ``` 
  to format a number `x` in decimal format according to style conventions found 
  in many European countries.
  
* provide a suitable labeller function to the `labels` argument of 
  `scale_xxx_yyy` so you end up with
  ```
  scale_x_continuous(
    labels = scales::label_percent(big.mark = ".", decimal.mark = ",")
  )
  ```
  to display numbers on the x axis as percentages using the style convention 
  found in many European countries.

## The solution

A first and simple solution to this problem would be to add some simple wrappers 
at the beginning of your R script or R markdown document like
```
my_label_number <- function() {
  scales::label_number(big.mark = ".", decimal.mark = ",")
}
```
But then you end up copy and pasting from one report to the next, so sooner or 
later you probably put these helpers inside package. 

And that's the goal of `countryscales`: Providing out-of-the-box helpers to 
format numbers using country-specific style conventions.

As a first example consider formatting a number `r x <- 12345689.987654321` 
according to style conventions used in Germany and several other European 
countries where a dot (`.`) is used as the big mark or grouping mark or 
thousands seperator and a comma (`,`) as the decimal mark.

Using `scales::label_number` this requires to switch the default decimal and big 
marks:
```{r}
library(scales)

label_number(big.mark = ".", decimal.mark = ",", accuracy = .1)(x)
```

Using `countryscales` this could be achieved with less typing using 
`countryscales::label_number_de`:

```{r}
library(countryscales)

label_number_de(accuracy = .1)(x)
```

To provide an example of using `countryscales` with `ggplot2` let's first 
prepare a small example dataset of the top 10 countries according to population 
size:

```{r top-10-pop}
top10_pop <- gapminder15[order(-gapminder15$pop), c("country", "pop")]
top10_pop$pct <- top10_pop$pop / sum(top10_pop$pop)
top10_pop <- head(top10_pop, 10)
```

As a basic example let's make a simple barchart of population size by country

```{r}
library(ggplot2)

p <- ggplot(top10_pop, aes(pop, reorder(country, pop))) +
  geom_col() +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(
    x = NULL, y = NULL,
    title = "Top 10 of countries by population size in 2015",
    caption = "Data: gapminder.org"
  )
```
While this chart is fine it's not ready for publication. Let's say we want to 
display population sizes on the x axis in decimal format and additionally add 
the share of each country on World population formatted as percentages as labels 
to the bars using German style conventions. 

Using `scales` this could be achieved like so:

```{style-scales}
p +
  # Add percentages to bars using German style conventions
  geom_text(
    aes(
      label = label_percent(big.mark = ".", 
                            decimal.mark = ",", 
                            accuracy = .1)(pct)), 
    hjust = 1.1, size = 8 / .pt, color = "white") +
  # Format numbers as percentages
  scale_x_continuous(
    labels = label_number(big.mark = ".", 
                          decimal.mark = ","), 
    expand = c(0, 0, .05, 0))
```

As this simple example shows displaying numbers in decimal format or as 
percentages requires some typing, especially if you want to or have to deviate 
from default decimal and grouping marks. 

That's where `countryscales` comes in handy as using `label_percent_de` and 
`scale_x_number_de` the same could be achieved with less typing like so:

```{r style-countryscales}
p +
  # Add percentages to bars using German style conventions
  geom_text(
    aes(
      label = label_percent_de(accuracy = .1)(pct)), 
    hjust = 1.1, size = 8 / .pt, color = "white") +
  # Format numbers as percentages
  scale_x_number_de(expand = c(0, 0, .05, 0))
```
